Exploring the Determinants of Body Weight: A Bayesian Analysis of Anthropometric and Demographic Factors Among Female U.S. Army Personnel

### 2024-12-04 paper/paper.qmd
rendered paper
added model comparison

All the data analysis was done through R (R Core Team 2023) with the aid of various packages in the R programming language. The `tidyverse` package [@tidyverse] was used to streamline the entire process of data manipulation, transformation, and visualization, ensuring consistency, efficiency, and flexibility throughout the analysis. Its comprehensive suite of tools enabled the creation of tailored and informative visual outputs, as well as the effective transformation and summarization of complex datasets.

@@

## Reflecting Data Features in Model Design

@@

## Justification for the Priors and Model Choice

@@

A frequentist linear regression model was also considered for its simplicity and interpretability; however, it does not provide posterior distributions or credible intervals, which are essential for quantifying uncertainty in this study. By contrast, the Bayesian regression framework provides probabilistic estimates, enabling uncertainty quantification and credible intervals for the coefficients. This approach aligns well with the study's objectives, where understanding the relative influence of predictors on body weight requires a robust and interpretable model. The rstanarm package was chosen for its efficient handling of Bayesian inference and built-in diagnostics, ensuring rigorous validation. The Gaussian likelihood function complements the continuous nature of the outcome variable (body weight), and the model structure strikes a balance between simplicity and capturing key relationships in the data.

>>>>

All the data analysis was done through `R` with the aid of various packages in the `R` programming language [@citeR]. The `tidyverse` package [@tidyverse] was used to streamline the entire process of data manipulation, transformation, and visualization, ensuring consistency, efficiency, and flexibility throughout the analysis. Its comprehensive suite of tools enabled the creation of tailored and informative visual outputs, as well as the effective transformation and summarization of complex datasets.

@@

## Predictors Choice

@@

@@

A Frequentist linear regression model was also initially considered for its simplicity and interpretability; however, it does not provide posterior distributions or credible intervals, which are essential for quantifying uncertainty in this study. By contrast, the Bayesian regression framework provides probabilistic estimates, enabling uncertainty quantification and credible intervals for the coefficients. This approach aligns well with the study's objectives, where understanding the relative influence of predictors on body weight requires a robust and interpretable model. The `rstanarm` package was chosen for its efficient handling of Bayesian inference and built-in diagnostics, ensuring rigorous validation. The Gaussian likelihood function complements the continuous nature of the outcome variable (body weight), and the model structure strikes a balance between simplicity and capturing key relationships in the data.

## Model Selection

The model selection process involved building and comparing two Bayesian regression models using the analysis data: a full model with all predictors and a reduced model with only age and component as predictors. The data was split into training (80%) and testing (20%) subsets for model evaluation. Both models were fit using the stan_glm function, and predictions were made on the test set. The models were evaluated using Root Mean Square Error (RMSE) as a metric of predictive performance. The results showed that the full model had a slightly lower RMSE (133.16) compared to the reduced model (134.49), indicating that the additional predictors in the full model improved its predictive accuracy, albeit marginally.
